\documentclass[]{phdthesis}
\addbibresource{../../references/thesis_n.bib}

\degree{Doctor of Philosophy}
\university{University College London}
\title{Computational design and\\[0.2em]characterisation of synthetic\\[0.2em] genetic switches}
\author{Miriam Leon}
\primarysupervisor{Dr. Chris P Barnes}
\secondarysupervisor{Prof. Geraint MH Thomas}

%%  Abbreviations entries  -----------------%%
%\acrlong{ } whole phrase, \acrshort{ } acronym, \acrfull{ } both
\makeglossaries
\input{../../glossary.tex}
\newcommand{\myrightleftarrows}[1]{\mathrel{\substack{\xrightarrow{#1} \\[-.9ex] \xleftarrow{#1}}}}

%%---------------------------------------------------%%
\begin{document}
% front matter
\maketitle


\begin{abstract}
Genetic toggle switches consist of two mutually repressing transcription factors. The switch motif forms the basis of epigenetic memory and is found in natural decision making systems, such as cell fate determination in developmental pathways. A synthetic genetic switch can be used for a variety of applications, like recording the presence of different environmental signals, for changing phenotype using synthetic inputs and as building blocks for higher-level sequential logic circuits. 

In this thesis, the genetic toggle switch was studied computationally and experimentally. Bayesian model selection methods were used to compare competing model designs of the genetic toggle switch. It was found that the addition of positive feedback loops to the genetic toggle switch increases the parametric robustness of the system.

A computational tool based on Bayesian statistics was developed, that can identify regions of parameter space capable of producing multistable behaviour while handling parameter and initial conditions uncertainty. A collection of models of genetic switches were examined, ranging from the deterministic simplified toggle switch to stochastic models containing different positive feedback connections. The design principles behind making a bistable switch were uncovered, as well as those necessary to make a tristable or quadristable switch.

Flow Cytometry was used to characterise a known toggle switch plasmid. A computational tool was developed which uses Bayesian statistics to infer model parameter values from flow cytometry data. This tool was used to characterise the toggle switch plasmid and fit a stochastic computational model to experimental data.

The work presented here suggests ways in which the construction of genetic switches can be enhanced. The algorithms developed were shown to be useful in synthetic system design as well as parameter inference. The tools developed here can enhance our understanding of biological systems and constitute an important addition to the systems approach to synthetic biology engineering. 


\end{abstract}
\begin{acknowledgements}

I would like to thank the people whose help and support made the work presented in this thesis possible.


First and foremost I would like to thank my supervisor, Dr. Chris Barnes. The hard work and dedication you show your students have always been inspiring appreciated. None of this would have been possible without your push, encouragement and patience which were sometimes the only things that kept me going. I was lucky to be your first student, a badge I will always wear with pride! Thank you for  trusting me, a young biologist, with this project and for always having the time and will to help me through this journey.

I would also like to thank Professor Geraint Thomas. Your cheerful attitude to science has never failed to lift my spirits. Your ability to put problems into perspective has always allowed me to keep pushing forward. 

I am grateful for the funding I received through the UCL Impact Award scheme.

Thank you to Professor David Hall for his help with plasmid cloning and flow cytometry. I would also like to thank Dr Gyorgy Szabadkai's lab for their help with flow cytometry, Dr. Ayad Eddaoudi in the flow cytometry core facility at the UCL lnstitute of Child Health and Dr Malte Paulsen at the Flow Cytometry Core Facility at the National Heart \& Lung Institute (NHLI), Imperial College London. 

To all the members of the Computational Systems and Synthetic Biology (CSSB) group, especially Alex Fedorec, Riana Gaifulina, David Gonzales, Aaran Lewis, Phil Lewis, Tanel Ozdemir, Lourdes Sriraja, Gordon Walsh, Marc Williams, Bethan Wolfenden, and Mae Woods. Thank you all for your constant support and friendship. You have made all the difference in making this journey an enjoyable one. A special thanks to Alex Fedorec for proofreading this thesis and to Tanel Ozdemir for spending hours teaching me how to work in the lab and to both for always being there to brighten the day. A special thanks also to Dr. Mae Woods for always helping me and being supportive with any maths and coding questions, for your kind words and for proofreading this thesis. I would also like to thank Dr. Robert Stanley for kindly providing the technical help in the writing of this thesis. And to all, thank you for being such a great team. 


To my friends that kept me sane throughout this journey, I thank all of you for your constant encouragement. To Effie and Eva for setting an example of what courage and hard work can achieve. To Daisy and Dafni for always believing in me and listening to my woes. To Daniel and Sinan for setting the bar high while encouraging me to reach it. To Adam, Jon, Meg, and Zosia for taking me on adventures when they were needed. 

To my family who have always been my cheerleaders. Mum, Dad, Evie thank you for your unconditional love and support. Σας αγαπώ πολύ. 

Finally, to Takao for your support and love. From going over algorithms to endless emotional support, your help in every aspect of this endeavour has made it all possible. 

\end{acknowledgements}
\tableofcontents*
\listoffigures
\listoftables
\glossarystyle{list}

\printglossary[type=\acronymtype, title=Abbreviations, toctitle=List of Abbreviations]

%\printglossary[type=\acronymtype, title=Abbreviations, toctitle=Abbreviations]



%% --------------------------------------------------------------
%% CHAPTERS
%% --------------------------------------------------------------

%% Introduction -------------------------------------%%
\mainmatter*
\chapter{Introduction}
\label{ch:Intro}
\input{../../chapters/chapterIntroduction/ChIntro_introduction}

%% Background -------------------------------------%%
\mainmatter*
\chapter{Background}
\label{ch:backg}
\input{../../chapters/chapterBackgr/ChBackground}


%% Switch literature -------------------------------------%%
%\mainmatter*
%\chapter{Current understanding of% the genetic toggle switch}
%\label{ch:swlit}
%\input{../../chapters/chapterLiterature/ChSwitch}

%% ABC-SysBio ---------------------------------------%%
\mainmatter*
\chapter{Positive feedback loops can increase the robustness of a genetic toggle switch}
\label{ch:abcsysbio}
\input{../../chapters/chapterABCSysBio/ChAbcSys_introduction}


%% StabilityFinder ----------------------------------%%
\mainmatter*
 
\chapter{Dynamics of multi-stable switches}
\label{ch:SF}
\input{../../chapters/chapterStabilityFinder/ChSF_introduction}
%\input{chapterStabilityFinder/methods}
%\input{chapterStabilityFinder/results}
%\input{chapterStabilityFinder/Lu_switches/lu_switches_summary}
%\input{chapterStabilityFinder/mass_action_switches/mass_action_switch_summary}
%\input{chapterStabilityFinder/discussion}


%%  Characterising the toggle switch -----------------%%
\mainmatter*
\chapter{Bayesian model fitting for flow cytometry data}
\label{ch:Flow}
\input{../../chapters/chapterABCFlow/ChFlow_and_exp_new_str}
%\input{../chapters/chapterCharacterisation/ChChar_introduction}
%\input{../chapters/chapterABCFlow/ChFlow_introduction}

%% Designing new switches ---------------------------%%
\mainmatter*
\chapter{Designing new switches}
\label{ch:desSw}
\input{../../chapters/chapterDesignSwitches/ChDesSw_introduction}

%% Chapter-6 Conclusions --------------------------------------%%
\mainmatter*
\chapter{Conclusions}
\label{ch:Conc}
\input{../../chapters/chapterConclusions/ChConc}

\printbibliography 

%% --------------------------------------------------------------
%% APPENDIX
%% --------------------------------------------------------------

%\addcontentsline{toc}{chapter}{Appendices}
\appendix
\chapter{Biochemical kinetic models}
\label{ap:ODEs}
\input{../../appendices/appendix_ODEs}

\chapter{Primers}
\label{ap:Prim}
\input{../../appendices/appendix_primers}

\chapter{Algorithms}
\label{ap:Alg}
\input{../../appendices/appendix_algos}

\chapter{Additional posterior distributions}
\label{ap:figures}
\input{../../appendices/appendix_posteriors}

\end{document}

